Constructing a universe for the setoid model

### 2020-07-06 paper.tex
typos

What we would like instead is a formulation of ITT that supports extensionality,
while retaining its convenient computational behaviour.
%
Unfortunately, one can show that canonicity and function extensionality cannot
coexist if propositional equality is given by Martin-L\"of's inductively-defined
identity types.
%
The first step towards a solution is to give up the idea of propositional
equality as a single inductive definition given homogeneously for an arbitrary
type. Instead, equality should be \emph{specific} to each type former in the
type theory, or in other words, every type former should be introduced alongside
an explanation of what counts as equality for its elements.
%% In this setting, function extensionality can simply be introduced as the
%% chosen notion of equality on function types.

This idea of pairing types together with their own equality relation goes back
to the notion of \emph{setoid} or \emph{Bishop set}. Setoids provide a quite
natural and useful semantic domain in which to interpret type theory. The first
setoid model was constructed to show consistency of funext for ITT
\cite{hofmann}. This consistency result is implied by the observation that UIP
and funext hold in the setoid model. Moreover, it was shown by Altenkirch
\cite{setoid99} that is the model construction is carried out in a type
theoretic metatheory with a universe of strict (definitionally proof-irrelevant)
propositions, it is possible to define a univalent universe of propositions
satisfying propositional extensionality. The setoid model thus satisfies all the
extensionality principles that we would like to have in type theory. The
question is whether there exists a version of intensional type theory that
supports setoid reasoning, and hence the forms of extensionality enabled by it.

>>>>

What we would like instead is a formulation of ITT that supports extensionality,
while retaining its convenient computational behaviour.
%
Unfortunately, one can show that canonicity and function extensionality cannot
coexist if propositional equality is given by Martin-L\"of's inductively-defined
identity types.
%
The first step towards a solution is to give up the idea of propositional
equality as a single inductive definition given homogeneously for arbitrary
types. Instead, equality should be \emph{specific} to each type former in the
type theory, or in other words, every type former should be introduced alongside
an explanation of what counts as equality for its elements.
%% In this setting, function extensionality can simply be introduced as the
%% chosen notion of equality on function types.

This idea of pairing types together with their own equality relation goes back
to the notion of \emph{setoid} or \emph{Bishop set}. Setoids provide a quite
natural and useful semantic domain in which to interpret type theory. The first
setoid model was constructed to show consistency of funext for ITT
\cite{hofmann}. This consistency result is implied by the observation that UIP
and funext hold in the setoid model. Moreover, it was shown by Altenkirch
\cite{setoid99} that if the model construction is carried out in a type
theoretic metatheory with a universe of strict (definitionally proof-irrelevant)
propositions, it is possible to define a univalent universe of propositions
satisfying propositional extensionality. The setoid model thus satisfies all the
extensionality principles that we would like to have in type theory. The
question is whether there exists a version of intensional type theory that
supports setoid reasoning, and hence the forms of extensionality enabled by it.
